---
title: "Dendrochronology in R"
author: "Jen Baron, j.baron@alumni.ubc.ca, UBC Tree Ring Lab"
date: "May 15, 2020"
output:
  html_document:
    keep_md: yes
    theme: flatly
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r global_options, echo = TRUE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

# Packages

```{r setup}
library(dplR) # dendrochronology
library(dplyr)
library(ggplot2)
```

# Load Data

```{r}
pinery.rwl <- read.rwl("data/pinery.rwl")
```


```{r}
master_all.rwl <- read.rwl("data/master_all.rwl")
master.rwl <- read.rwl("data/master_all.rwl")
mi015.rwl <- read.rwl("data/mi015.rwl")
cana127.rwl <- read.rwl("data/cana127.rwl")
cana148.rwl <- read.rwl("data/cana148.rwl")
pa008.rwl <- read.rwl("data/pa008.rwl")
```

```{r}
rwl.report(pinery.rwl)
```

## Basic Summary Statistics

```{r}
rwl.stats(pinery.rwl)
```

```{r}
seg.plot(pinery.rwl)
```

## Truncate Based on Date / Sample Depth

Truncate to 1898 (sample depth = 8)

```{r}
pinery.rwl.trunc <- pinery.rwl[46:166,]
#write.rwl(pinery.rwl.trunc, "outputs/pinery_rwl_trunc.rwl", format=c("tucson"))
```



# Crossdating


```{r}
pinery.corr <- corr.rwl.seg(pinery.rwl.trunc, seg.length = 50, 1918)
```

```{r}
#identify flags
flags <- as.data.frame(pinery.corr$flags) 
flags

#correlation by series
overall <- pinery.corr$overall
overall %>% round(2)


#average series correlation
avg.seg.rho <- as.data.frame(pinery.corr$avg.seg.rho)
avg.seg.rho %>% round(2)


#correlation by segment and series
rho <- as.data.frame(pinery.corr$spearman.rho)
rho %>% round(2)
```

There is (at least) one other way of looking at the average interseries
correlation of a data set. The interseries.cor function in dplR gives
a measure of average interseries correlation that is different from the rbar
measurements from rwi.stats. In this function, correlations are calculated
serially between each tree-ring series and a master chronology built from
all the other series in the rwl object (leave-one-out principle).

```{r}
interseries.cor(pinery.rwl) %>% round(2)
```


# Compare to Master Chronologies 

See Master Chronology .Rmd for details


## Compare to Filtered Regional Chronology

```{r}
master.corr <- corr.rwl.seg(pinery.rwl.trunc, master=master.rwl,seg.length=30, 1853)
```

```{r}
#identify flags
flags.2 <- as.data.frame(master.corr$flags) 
flags.2

#correlation by series
overall.2 <- master.corr$overall %>% as.data.frame()
overall.2 %>% round(2)


#average series correlation
avg.seg.rho.2 <- as.data.frame(master.corr$avg.seg.rho)
avg.seg.rho.2 %>% round(2)


#correlation by segment and series
rho.2 <- as.data.frame(master.corr$spearman.rho)
rho.2 %>% round(2)


summarize(overall.2, mean = mean(rho)) %>% round(2)
```


## Compare to Full Regional Chronology

```{r}
master.corr.all <- corr.rwl.seg(pinery.rwl.trunc, master=master_all.rwl, seg.length=50, 1853)
```

```{r}
#identify flags
flags.3 <- as.data.frame(master.corr.all$flags) 
flags.3

#correlation by series
overall.3 <- master.corr.all$overall %>% as.data.frame()
overall.3 %>% round(2)


#average series correlation
avg.seg.rho.3 <- as.data.frame(master.corr.all$avg.seg.rho)
avg.seg.rho.3 %>% round(2)


#correlation by segment and series
rho.3 <- as.data.frame(master.corr.all$spearman.rho)
rho.3 %>% round(2)


summarize(overall.3, mean = mean(rho)) %>% round(2)
```


## Compare to Each Master Chronlogy

### Wyse - Platte Plains - PIST - ITRDB MI015


```{r}
rwl.report(mi015.rwl)
master.corr.mi015 <- corr.rwl.seg(pinery.rwl.trunc, master=mi015.rwl, seg.length = 30, bin.floor = 1913)
```

```{r}
#identify flags
flags.4 <- as.data.frame(master.corr.mi015$flags) 
flags.4

#correlation by series
overall.4 <- master.corr.mi015$overall %>% as.data.frame()
overall.4 %>% round(2)


#average series correlation
avg.seg.rho.4 <- as.data.frame(master.corr.mi015$avg.seg.rho)
avg.seg.rho.4 %>% round(2)


#correlation by segment and series
rho.4 <- as.data.frame(master.corr.mi015$spearman.rho)
rho.4 %>% round(2)


summarize(overall.4, mean = mean(rho)) %>% round(2)
```

### Guyette - Dividing Lake - PIST - ITRDB CANA127

```{r}
rwl.report(cana127.rwl)
master.corr.cana127 <- corr.rwl.seg(pinery.rwl.trunc, master=cana127.rwl, seg.length = 30, bin.floor=1905)
```

```{r}
#identify flags
flags.5 <- as.data.frame(master.corr.cana127$flags) 
flags.5

#correlation by series
overall.5 <- master.corr.cana127$overall %>% as.data.frame()
overall.5 %>% round(2)


#average series correlation
avg.seg.rho.5 <- as.data.frame(master.corr.cana127$avg.seg.rho)
avg.seg.rho.5 %>% round(2)


#correlation by segment and series
rho.5 <- as.data.frame(master.corr.cana127$spearman.rho)
rho.5 %>% round(2)


summarize(overall.5, mean = mean(rho)) %>% round(2)
```

### Guyette - Dividing Lake Aquatic - PIST - ITRDB CANA148

```{r}
rwl.report(cana148.rwl)
master.corr.cana148 <- corr.rwl.seg(pinery.rwl.trunc, master=cana148.rwl, seg.length = 30, bin.floor=1904)
```

```{r}
#identify flags
flags.6 <- as.data.frame(master.corr.cana148$flags) 
flags.6

#correlation by series
overall.6 <- master.corr.cana148$overall %>% as.data.frame()
overall.6 %>% round(2)


#average series correlation
avg.seg.rho.6 <- as.data.frame(master.corr.cana148$avg.seg.rho) %>% as.data.frame
avg.seg.rho.6 %>% round(2)


#correlation by segment and series
rho.6 <- as.data.frame(master.corr.cana148$spearman.rho)
rho.6 %>% round(2)


summarize(overall.6, mean = mean(rho)) %>% round(2)
```


### Cook - Longfellow Trail - PIST - ITRDB PA008


```{r}
rwl.report(pa008.rwl)
master.corr.pa008 <- corr.rwl.seg(pinery.rwl.trunc, master=pa008.rwl, seg.length=40, bin.floor = 1902)
```



```{r}
#identify flags
flags.7 <- as.data.frame(master.corr.pa008$flags) 
flags.7

#correlation by series
overall.7 <- master.corr.pa008$overall %>% as.data.frame()
overall.7 %>% round(2)


#average series correlation
avg.seg.rho.7 <- as.data.frame(master.corr.pa008$avg.seg.rho)
avg.seg.rho.7 %>% round(2)


#correlation by segment and series
rho.7 <- as.data.frame(master.corr.pa008$spearman.rho)
rho.7 %>% round(2)


summarize(overall.7, mean = mean(rho)) %>% round(2)
```














# Detrending

## Compare Methods

```{r}
detrend(rwl = pinery.rwl.trunc, method = c("ModNegExp", "Spline"), make.plot = TRUE, verbose=TRUE, return.info = TRUE)
```



## Apply Detrending Method
```{r}
pinery.rwi.info <- detrend(rwl = pinery.rwl.trunc, method = c("ModNegExp"), return.info = TRUE)

pinery.rwi <- pinery.rwi.info$series
```




## Descriptive Statistics

### Read IDs

```{r}
pinery.ids <- read.ids(pinery.rwl.trunc)
pinery.ids
```

### Standard Chronology

```{r}
rwi.stats(pinery.rwi, pinery.ids, prewhiten=FALSE) #running.window=TRUE, window.length = 50)
```


### Residual Chronology

```{r}
rwi.stats(pinery.rwi, pinery.ids, prewhiten=TRUE) #running.window=TRUE, window.length = 50)
```


### Interseries Correlations

```{r}
interseries.cor(pinery.rwi, biweight=TRUE) %>% round(3)
```


# Building a Mean Value Chronology

The simplest way to make a chronology in dplR is chronology is with the crn function which also has a plot method. This defaults to building a mean-value chronology by averaging the rows of the rwi data using Tukey’s biweight robust mean (function tbrm in dplR). 


```{r}
pinery.crn <- chron(pinery.rwi, prewhiten=TRUE, biweight = TRUE)
head(pinery.crn)

pinery.crn$samp.depth <- NULL
```

## Plot Chronology

```{r}
crn.plot(pinery.crn, add.spline = TRUE, nyrs=15)
```


## Add Uncertainty Estimates

```{r}
pinery.avg <- apply(pinery.rwi,1,mean,na.rm=TRUE)

yrs <- time(pinery.crn)


se <- function(x){
  x2 <- na.omit(x)
  n <- length(x2)
  sd(x2)/sqrt(n)}


pinery.se <- apply(pinery.rwi,1,se)

pinery.sd <- apply(pinery.rwi,1,sd,na.rm=TRUE)

par(mar = c(2, 2, 2, 2), mgp = c(1.1, 0.1, 0), tcl = 0.25, xaxs='i')

plot(yrs, pinery.avg, type = "n", xlab = "Year", ylab = "RWI", axes=FALSE)
xx <- c(yrs,rev(yrs))
yy <- c(pinery.avg+pinery.se*2,rev(pinery.avg-pinery.se*2))
polygon(xx,yy,col="grey80",border = NA)
lines(yrs, pinery.avg, col = "black")
lines(yrs, ffcsaps(pinery.avg, nyrs = 15), col = "red", lwd = 1) #adds a spline
axis(1); axis(2); axis(3); axis(4)
box()
```


# Conclude Session

```{r}
sessionInfo()
```

